The most common use for blockchains is in financial applications. This gives a crucial importance to the
security of the consensus protocol used in a blockchain. Unfortunately, many currently used blockchains are
vulnerable to a certain type of consensus attack, known as the \emph{bribery attack}. In a bribery attack, an adversary
tries to corrupt participants of a protocol by offering them money and seducing them to violate the protocol.

At the time of writing of this document,
the total mining reward for a Bitcoin block is around \$150,000, and the Bitcoin network approximately produces 5 blocks
per hour. If we assume, in decision theoretic terminology, that the mining reward solely defines the utility
function\footnote{In decision theory, the utility function measures the preference, worth or value of different
alternatives for a decision maker.} of a Bitcoin miner, one should be able to hire all hashing power
of the Bitcoin network for one hour by
spending only \$750,000. The situation is not much different for PoS blockchains, as long as the total
stake\footnote{Here by stake, we mean a real number measuring the total interest of a user in the system, and we are
not referring, in particular, to some locked amount of a user's money that is known as stake in some PoS protocols.}
of the validator set is a relatively small value.
This problem is even more severe in blockchains that use randomly selected small sets of validators. These small sets
usually have low total stake and could be easily bribed and corrupted. Selecting these random sets
by hidden random procedures would not help, since the validator himself knows he has been selected before casting
his vote.

It appears that the only solution to this important vulnerability is to effectively participate all the stakeholders
of the system in the consensus protocol. This large participation makes the protocol resilient against bribery or
collusion because the adversary would need to spend unrealistic amounts of money to bribe enough users.

However, for effective participation in the consensus protocol, a validator needs to be able to detect
illegal transactions. Detecting illegal transactions can be done by accessing the ledger
state and executing transactions according to the protocol rules. The
ledger state, even for small blockchains, could be several
hundreds of gigabytes, and executing transactions could easily become costly when a blockchain is acting as a
smart contract platform. This computational and storage overhead, in practice, could prevent most of
ordinary users from any type of participation in the consensus protocol of a blockchain.

A fully decentralized blockchain based on the participation of every user looks appealing, though it is not as perfect
as it might seem. The consensus protocol of a blockchain relies on a network of computers, not humans. Ordinary users
use simple and similar computer systems. That means, they all have similar vulnerabilities and weaknesses which could be
used by an adversary to catastrophically attack the consensus protocol. For instance, if a malware, probably using a
common zero-day
vulnerability, has the ability to infect a large portion of normal personal computers, an adversary can use it to
take control of the majority of participants in the consensus protocol and compromise the security of the
blockchain.

Securing a computer system against cyberattacks needs planning ahead and access to engineering resources.
Special software and hardware, like custom-built operating systems and isolated specialized hardware are required.
This is not something a normal user can afford. Powerful centralized entities, having large financial and
technical resources, could build
systems that are resilient against sophisticated cyberattacks. In this regard, we have to admit, a centralized system is
arguably superior to a decentralized\footnote{Note that decentralized and distributed are two different concepts.}
system.

To overcome these difficulties, Argennon\footnote{The classical pronunciation should be used:\textipa{/Ar"gen.non/}}
uses a Hybrid Proof of Stake consensus protocol, which is
capable of combining the benefits of a centralized and a decentralized system. A small committee of
delegates is democratically elected by users via the Argennon Decentralized Autonomous Governance system
(ADAGs). This committee usually\footnote{The election term and the number of committee members can be changed by the
ADAGs.} is elected for a one-year term, has five members, and is responsible for minting new
blocks of the Argennon blockchain.
Each minted block gets certified by all members of the delegates committee and after that it must get approved
by its corresponding validator assembly.

Validator assemblies are very
large sets of validators including at least three percent of the maximum possible stake of the Argennon network.
Every validator assembly has an index between $0$ and $m - 1$, and is responsible for validating block number $n$,
if $n$ modulo $m$ equals the index of the assembly. A block is approved if it takes approval votes from at
least $2/3$ of the total stake of its validator assembly.

In case the main committee fails to generate new blocks or behaves maliciously, a special \emph{recovery protocol} is
initiated by validators.
This recovery protocol can recover the functionality
of the Argennon blockchain as long as more than $2/3$ of the total stake of every validator assembly is controlled by
honest users and any network partition resolves after a finite amount of time. The recovery protocol uses two main
emergency procedures to recover the functionality of the Argennon blockchain: \emph{Emergency Forking} and
\emph{Emergency Agreement} protocol.

In Argennon, a block is considered final after its \textbf{next} block gets certified by \textbf{both} the
delegates committee and its validator assembly.
The Argennon protocol ensures that as long as more than $2/3$ of the total stake of validators is
controlled by honest users, the probability of discarding a final block is near zero even if all the delegates are
malicious.

Each block of the Argennon blockchain contains a set of Computational Integrity (CI) statements and a commitment to
the final ledger state of the block (i.e.\ the ledger state after executing all transactions of the block). The CI
set defines how the final state of the block can be reached from
the state of its previous block via a set of intermediate state transitions. Each individual CI statement defines
an ordered list of external requests\footnote{External requests in Argennon are
similar to transactions in older blockchains.} and determines the state before and after executing those requests.

More formally, each CI statement has the form
$\hat{\mathbb{S}} \coloneqq \tau(\mathbb{S}, \mathbb{R})$, and states that
$\hat{\mathbb{S}}$ is the commitment of the next state after executing an ordered list of external requests with
commitment $\mathbb{R}$ on a state which has commitment $\mathbb{S}$.
Commitments to states are produced by a cryptographic accumulator. $\tau$ is a transition function and is known to
both the validator and the prover.

Validators can\footnote{Using the Argennon cloud is optional for a validator.} receive succinct cryptographic proofs
(STARK/SNARK) of these CI statements from the Argennon cloud and validate blocks without storing the ledger
state or performing costly computation.

Verifying a succinct proof can be exponentially faster than replaying the computation. Moreover, verifications of
different CI statements are independent of each other and can be done in parallel. As a result, a validator can use
multiple cores for verifying CI statements of a \textbf{single block} and different validator assemblies can
simultaneously and
independently verify \textbf{different blocks}.
In addition, proof generation of different CI statements similarly can be done in parallel. However, for parallel
proof generation, the state transition needs to be known in advance. That's why in Argennon, delegates do not try
to generate proofs and focus all their computational power on executing external requests and
generating the state transition as fast as possible.

On the blockchain, Argennon applications (i.e smart contracts) are stored in a high level text based language, called
the Argennon Standard Application Representation (ASAR). The ASAR is intended for preserving the
high level information of the application logic to facilitate platform specific compiler optimization at a host
machine. This enables delegates to compile and optimize Argennon applications for their specific hardware platforms
and execute applications efficiently, ensuring the state transition can be found as fast as possible.

In addition to the ASAR, an Argennon application is stored on the blockchain in the Argennon Prover Machine (APM) code.
This code is used by validators for CI proof verifications. The Argennon Prover Machine is a virtual
machine tailored for efficient verification of AscEE computations by
argument systems. The APM has a minimal RISC architecture with a very compact instruction set. This ensures that
its transition function has an efficient circuit complexity. However, the Argennon protocol does not enforce the usage
of the APM. Validators and PVC servers can use any argument system with any arithmetization, and if required, the ASAR
of an application can be used for generating the appropriate arithmetization instead of the APM\@.

Independence of CI statements is useful, but is not enough for having a truly scalable blockchain. To increase
parallelism, the Argennon protocol enforces all external requests to pre-declare their memory access locations. That
would enable a block proposer\footnote{In Argennon, delegates are the only block proposers.} to use Data Dependency
Analysis\footnote{See Section~\ref{sec:concurrency}} (DDA) to indicate independent sets of external requests (i.e.\
transactions) and use those sets for parallel processing. More importantly, these sets can be used for generating CI
statements that are defined on the \textbf{same initial state} and their proof can be generated independently without
the need to calculate the state transition in advance.

Centralized block generation brings some interesting features to the Argennon platform, such as flexible and lower
fees, off chain fee payment, optimistic instant transaction confirmation, and front running protection.
However, it also increases the possibility of transaction censorship. In Argennon, this problem is addressed by a
special High Priority Request (HPR) protocol.

Using Succinct Argument of Knowledge systems makes the main functionalities of an Argennon validator light enough to
be implemented as a smart contract. By deploying a validator contract on another platform, Argennon
could use more established blockchains as an extra layer of security, specially
during the bootstrapping phase, when ARG is not well distributed yet. In addition, this contract will facilitate
trustless bridging of assets from that platform to the Argennon blockchain. To reduce execution fee, only
roll-ups of the state transition can be validated by the contract.

The Argennon protocol strongly incentivizes the formation of a \textbf{permission-less} network of Publicly Verifiable
Cloud (PVC) servers. To do so, the Argennon protocol conducts repetitive automatic lotteries between PVC servers.
A PVC server can increase its chance of winning by (i) generating proofs for more CI statements, (ii) storing all
parts of the ledger state and providing \emph{proofs of storage}.

A PVC server in Argennon, is a conventional data server which uses its computational and
storage resources to help the Argennon network process transactions. This encourages the development
of conventional networking, storage and compute hardware, which can benefit all areas of information technology.
This contrasts with the approach of some older blockchains that incentivized the development of a totally
useless technology of hash calculation.
